---
title: "An Introduction to R"
author: "Pablo E. Gutiérrez-Fonseca"
date: "8/28/2021"
output: pdf_document
---

What is R?
Is a computer language that allows the user to program algorithms and use tools that have been programmed by others. 

R is based on the computer language S, which was developed by John Chambers and others at Bell Laboratories in 1976. In the early 1990s, Ross
Ihaka and Robert Gentleman (University of Auckland, in Auckland, New Zealand) experimented with the language and called their product R. Note
that both their first names begin with the letter R. Also when modifying a
language called S, it is tempting to call it T or R.

R comes with a series of default packages. A package is a collection of previously programmed functions, often including functions for specific tasks. It is tempting to call this a library, but the R community refers to it as a package.

" > " Significa "What now?"
The screen prompt > is an invitation to put R to work.
Each line can have at most 128 characters,   When continuation is expected, the prompt changes from > to +
Note that the + continuation prompt does not carry out arithmetic plus.

1.  Primer paso para comenzar a trabajar en R es instalar y llamar las librerias con las que vamos a trabajar.  
En cualquier momento se pueden llamar, pero si las pones arriba sera mejor visualizarlas luego.  

```{r}
# install.packages("Kendall")

library(Kendall)
library(tibble)

citation('Kendall')

```

1.1  Si necesitas ayuda de como trabaja una funcion en R puedes usar "  ? "

```{r}

# ?lm

```


2. Si comienzas un analisis nuevo.  Recomiendo limpiar en "Environment".

```{r}

# rm(list = ls(all = TRUE))

```


3. Hay que definir el directorio donde estas trabajando.
A no ser que estes en un proyecto en Github donde la documentacion (scripts, RMarkdown) y las tablas estan juntas.

```{r}

# setwd('D:/Curriculum/14_ Colaboracion/2021 Pulse LTER/Pulse WG PR/PulseDynWorkingGroup_Luq/data')

```


4. Importar tablas a R.  Hay varias maneras.

```{r}

# xxx.frm=read.table("xxx.txt",header=T)
# attach(xxx.frm)
#xxx.frm


# xxx.frm=read.csv("xxx.csv")
# attach(xxx.frm)
# xxx.frm


```

4.1 Se puede importar tablas seleccionandalas directamente

```{r}

# data<-read.table(file.choose(),header=T)

```


5. Construccion de tablas en R.

Usando data.frame

```{r}
    example<- data.frame(Day=c(2,4,8,16,32,44,2,4,8,16,32,44,2,4,8,16,32,44),
          Replicate=c(1,1,1,1,1,1,2,2,2,2,2,2,3,3,3,3,3,3,
                      1,1,1,1,1,1,2,2,2,2,2,2,3,3,3,3,3,3,
                      1,1,1,1,1,1,2,2,2,2,2,2,3,3,3,3,3,3),
          Treament=c("CC","CC","CC","CC","CC","CC","CC","CC","CC","CC","CC","CC","CC","CC","CC","CC","CC","CC",
                     "HP","HP","HP","HP","HP","HP","HP","HP","HP","HP","HP","HP","HP","HP","HP","HP","HP","HP",
                     "LL","LL","LL","LL","LL","LL","LL","LL","LL","LL","LL","LL","LL","LL","LL","LL","LL","LL"),
          AFDM=c(94.669342,94.465752,84.897023,81.435993,86.556221,75.328294,94.262162,88.791240,75.735474,81.232403,
                 67.050593,76.346244,95.076522,88.968823,83.879073,73.958836,70.645724,67.184695,99.763156,92.022673,
                 92.245362,74.513934,50.083136,36.979418,94.872932,86.353037,81.843173,67.795465,46.622106,18.323099,
                 95.089932,93.244212,81.679814,65.352385,18.286525,7.517794,99.559972,86.759404,84.693433,79.196504,
                 67.456961,54.765706,94.074014,87.543693,82.492548,72.333367,51.304676,51.304676,98.340870,86.322153,
                 87.950873,84.693433,63.316485,63.723665))

example


```

 6.1 Usando tribble
 
```{r}

 data <- tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)
                 
data
                 
```


7. Simbolos comunes en r, que vamos a estar usando.

```{r, eval = FALSE}
%>% dplyr. What the function does is to pass the left hand side of the operator to the first argument of the right hand side of the operator.
  
$ allows you extract elements by name from a named list. For example

```
