---
title: "nMDS"
author: "Pablo E. Gutiérrez-Fonseca"
date: "9/25/2021"
output: pdf_document
---

```{r, warning=FALSE,message=FALSE}

libraries <- c("vegan", "ggplot2", "dplyr")
lapply(libraries, require, character.only = TRUE)

```

Paso 1. LLamar a la tabla (como .csv)

```{r, results='hide'}
Moth_full  <- read.csv("data/fullmatrix.csv")
head(Moth_full)

```

Paso 2.  Seleccionar las especies.

```{r, results='hide'}
moth_sp <- dplyr::select(Moth_full, M1:A248)
str(moth_sp)
ncol(moth_sp)
nrow(moth_sp)

```
Paso 3.  Vamos hacer el nMDS

As a rule of thumb literature has identified the following cut-off values for stress-level:

Higher than 0.2 is poor (risks for false interpretation).
0.1 - 0.2 is fair (some distances can be misleading for interpretation).
0.05 - 0.1 is good (can be confident in inferences from plot).
Less than 0.05 is excellent (this can be rare).

An important number to note is the stress, which is roughly the “goodness of fit” of your NMDS ordination. For a good representation of your data, the stress value should ideally be less than 0.2. 

```{r}
set.seed(1) # Con este comando, siempre comenzara del mismo lugar.

moth.mds <- metaMDS(moth_sp, distance = "bray", k = 2, trymax=100)  #using all the defaults
moth.mds

```


Paso 4.  Vamos hacer el grafico. Un muy sensillo Plot

```{r}

plot(moth.mds, type="t")

```


Paso 5. Mirar el stressplot
Large scatter around the line suggests that original dissimilarities are not well preserved in the reduced number of dimensions. Looks pretty good in this case.

```{r}

stressplot(moth.mds)

```


Paso 6.  Otra metrica para evaluar el ajuste (buen desempeno) del nMDS es mirar el "Goodness of fit".

Visitar esta pagina para mas informacion:
https://rdrr.io/rforge/vegan/man/goodness.metaMDS.html

```{r}
gof <- goodness(moth.mds)
gof

{plot (moth.mds, display = 'sites', type = 't', main = 'Goodness of fit') # this function draws NMDS ordination diagram with sites
points (moth.mds, display = 'sites', cex = 2*gof/mean(gof))} # and this adds the points with size reflecting goodness of fit (bigger = worse fit)

```


Paso 7.  Chequear el % variacion explicado por cada Axis.

```{r}
MDS <- cmdscale(vegdist(moth_sp, method = "bray"), k = 2, eig = T, add = T )
round(MDS$eig*100/sum(MDS$eig),1)
      
```


Paso 8.  Cuantas dimenciones neceita para reducir el stress.

To decide how much dimensions you need you might plot the stress as a function of the number of dimensions. Note that, in contrast with a traditional scree plot, each bar does not represent the variance associated with each axis but the total stress (a function of the squared difference between d and d_hat) for all the dimensions. For example, the "3Dim" bar represent the stress of a solution in 3 dimensions, not the stress associated with the 3rd axis...
Here the improvements of the representation for dimensions > than 2 dimensions are low.

#  https://stackoverflow.com/questions/49223740/cumulative-variance-explained-for-nmds-in-r

```{r}
n = 10
stress <- vector(length = n)
for (i in 1:n) {
    stress[i] <- metaMDS(moth_sp, distance = "bray", k = i)$stress
}
names(stress) <- paste0(1:n, "Dim")
# x11(width = 10/2.54, height = 7/2.54)
par(mar = c(3.5,3.5,1,1), mgp = c(2, 0.6, 0), cex = 0.8, las = 2)

barplot(stress, ylab = "stress", main="NMDS Stepdown Stress Plot")


```



Paso 9.  Vamos a poner Habitat, Site y Periodo en el grafico. 
Primer las busco en la Matrix Orignal 

```{r}
Habitat <- select(Moth_full, Habitat)
Site <- select(Moth_full, Site)
Period <- select(Moth_full, Period)

```

Paso 10.  Extraer las cordenadas de los Axis del nmds.

```{r}

data.scores <- as.data.frame(scores(moth.mds))  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores

```

Paso 11.  Unir los datos de Habitat, Site y Periodo a mi nueva dataframe (i.e., data.scores)

```{r}

data.scores$Site <- unlist(Site)  #   create a column of site names
data.scores$Period <- unlist(Period)  
data.scores$Habitat <- unlist(Habitat)
head(data.scores) #look at the data
```

Paso 12.  Extraer las cordenadas de las especies y ponerlas en una nueva matrix.

```{r}
species.scores <- as.data.frame(scores(moth.mds, "species"))
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores

```


Paso 13.  Vamos hacer graficos bonitos en ggplot2.
Solo Especies.


```{r}

p <- ggplot() + 
  geom_text(data=species.scores, aes(x=NMDS1,y=NMDS2, label = "*"),size=7, alpha=0.5) # add the species labels

p

```

Paso 14.  Vamos a poner etiquetas. En este caso solo Habitats y Periordo

```{r}
p1 <- p + geom_point(data=data.scores,aes(x=NMDS1,y=NMDS2,
                                          shape=Period,colour=Habitat),size=5) + # add the point markers
        scale_color_manual(values=c("#00AFBB", "#FC4E07")) + 
        scale_shape_manual(values=c(15, 17))
p1 

```

Paso 15.  Agrerar sitios.

```{r}
p2 <- p1 + geom_text(data=data.scores,aes(x=NMDS1,y=NMDS2,label=""),size=6,vjust=0)  # add the site labels
p2 

```


Paso 16.  Graficos de Publicacion.

```{r}
p3 <- p1 + coord_equal() +
  theme_bw() + 
  theme(legend.title = element_text(size=12),
        axis.text.x = element_blank(),  # remove x-axis text
        axis.text.y = element_blank(), # remove y-axis text
        axis.ticks = element_blank(),  # remove axis ticks
        axis.title.x = element_text(size=16), # remove x-axis labels
        axis.title.y = element_text(size=16), # remove y-axis labels
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_blank()) + 
  guides(color = guide_legend(override.aes = list(shape = 16, size = 4)))
p3

```



