---
title: "partial RDA"
author: "Pablo E. Gutiérrez-Fonseca"
date: "11/17/2021"
output: html_document
---


1. Primer paso: cargar las librerias que necesitas.

```{r, warning=FALSE, message=FALSE}

library(ggplot2)
library(dplyr)
library(vegan)

```



2. Segundo paso: cargar los datos.

```{r}

species=read.csv("data/RDA_species.csv", header=T, row.names=NULL, sep=",")
env=read.csv("data/RDA_environmetal_standart.csv", header=T, row.names=NULL, sep=",")

```

<br/>

3. Remover la columna de sitos.
```{r}
species_1 <- select(species, -site)
env_1 <- select(env, -site)

```

<br/>

4.  Transformar datos.
Hellinger es una transformacion recomendada por Legendre & Callagher (2001)
en datos de abundancia y con una respuesta lineal
```{r}

species_2 <- decostand(species_1, method = "hellinger")

```


4. Vamos a construir los modelos: combinados e individuales
```{r}
partial_rda_all = rda(species_2 ~., data= env_1, scale=F)
partial_rda_env = rda(species_2 ~ temperature + oxygen + pH + conductivity, 
                      data= env_1, scale=F)
partial_rda_phy = rda(species_2 ~ plants + land_use + margin, 
                      data= env_1, scale=F)
partial_rda_season = rda(species_2 ~ season, data= env_1, scale=F)

```

5. Ver los R2adj de cada uno

```{r}
RsquareAdj (partial_rda_all)
RsquareAdj (partial_rda_env)
RsquareAdj(partial_rda_phy)
RsquareAdj(partial_rda_season)

```


```{r}
all <- RsquareAdj (partial_rda_all)$adj.r.squared
env <- RsquareAdj (partial_rda_env)$adj.r.squared
phy <- RsquareAdj (partial_rda_phy)$adj.r.squared
seas <- RsquareAdj(partial_rda_season)$adj.r.squared

a <- env + phy + seas - all  # 0.02125216
b <- env - a         # 0.03988225
c <- phy - a         # 0.06750099
d <- season - a         # 0.06750099

a;b;c;d

```

<br/>

6. Vamos a correr un RDA con las matrices (diferentes factores) separadas. 
```{r}
varp <- varpart(species_2, ~ temperature + oxygen + pH + conductivity, 
                ~ plants + land_use + margin, ~season, data = env_1)

varp
```


<br/>

Plot the results
The plot shows the adjusted R2 values associated with each partition or for overlapping partitions.

```{r}

plot(varp, digits = 2, Xnames = c('Enviromental factors', 'Plants', "Season"), 
     bg = c('navy', 'tomato', "lightblue"))

## Show values for all partitions by putting 'cutoff' low enough:
plot(varp, cutoff = -Inf, cex = 0.7, bg=2:5)
```

<br/>

Then we can test the significance of each individual component. 
```{r}

# significance of partition from Environment (physicochemical variables)
anova(rda(species_2 ~ temperature + oxygen + pH + conductivity + 
          Condition(plants + land_use + margin) +  
          Condition(season), data=env_1))

# Significance of partition from Physical characteristics
anova(rda(species_2 ~ plants + land_use + margin +
          Condition(temperature + oxygen + pH + conductivity) +  
          Condition(season), data=env_1))

# Significance of partition from Season
anova(rda(species_2 ~ season +
          Condition(plants + land_use + margin) +
          Condition(temperature + oxygen + pH + conductivity), 
          data=env_1))

```

Physicochemical and Physical characteriscs are all statistically significant in their contributions to Odonata community composition, even though the amount of variation explained by Physicochemistry is small. There was no statistically significant of season in Odonata composition independent of these other measured drivers.
